Text Is All You Need

### 2023-05-05 docs/hes_final/article.qmd
Two typo corrections

Predicting conflict has long been of interest to researchers and practitioners of international relations, and statistical and computational methods to do so can be found in the literature as early as 1944. Modern methodologies that resemble those we outline here began to appear in the literature in the early 2010s, when researchers began using holdout datasets for model validation [@10.2307/40730711]. This shift followed work by Ward et al that sharply criticized the use of statistical significance tests on in-sample prediction to validate conflict models [@doi:10.1177/0022343309356491]. Our methodology closely mirrors that which emerged in the following decade, including that described in more contemporary work, such as Mueller and Ruah [@RePEc:cam:camdae:2220]. In essence, this methodology involves using time-lagged input data to predict a conflict label that corresponds to a date some time after the input data would have become available. The model is then validated using an unseen holdout set, typically at the end of the date range available in the data [@doi:10.1080/03050629.2022.2070745]. This overcomes the problem of overfitting observed in earlier methods, and provides a principled starting point from which different models can be compared.

@@

### Defining Conflict Escalation
Our task of modeling conflict escalation required a precise definition of what conflict escalation means. We began with raw event and casualty counts from the [Armed Conflict Location & Event Data Project (ACLED)](https://acleddata.com/), who's team of researchers monitor thousands of sources in 20 languages to generate a dataset containing monthly totals of conflict events and their related fatalities [@doi:10.1177/0022343310378914]. ACLED defines these events as incidents involving armed groups with political objectives as well as demonstrations, riots, and protests. With these totals in-hand, we needed to create a target to label our training data. Prior work has treated this as a binary classification task, asking whether or not conflict escalated in a given month, and we adopt this approach as well. [@10.1093/jeea/jvac025]

>>>>

Predicting conflict has long been of interest to researchers and practitioners of international relations, and statistical and computational methods to do so can be found in the literature as early as 1944. Modern methodologies that resemble those we outline here began to appear in the literature in the early 2010s, when researchers began using holdout datasets for model validation [@10.2307/40730711]. This shift followed work by Ward et al. that sharply criticized the use of statistical significance tests on in-sample prediction to validate conflict models [@doi:10.1177/0022343309356491]. Our methodology closely mirrors that which emerged in the following decade, including that described in more contemporary work, such as Mueller and Ruah [@RePEc:cam:camdae:2220]. In essence, this methodology involves using time-lagged input data to predict a conflict label that corresponds to a date some time after the input data would have become available. The model is then validated using an unseen holdout set, typically at the end of the date range available in the data [@doi:10.1080/03050629.2022.2070745]. This overcomes the problem of overfitting observed in earlier methods, and provides a principled starting point from which different models can be compared.

@@

### Defining Conflict Escalation
Our task of modeling conflict escalation required a precise definition of what conflict escalation means. We began with raw event and casualty counts from the [Armed Conflict Location & Event Data Project (ACLED)](https://acleddata.com/), whose team of researchers monitor thousands of sources in 20 languages to generate a dataset containing monthly totals of conflict events and their related fatalities [@doi:10.1177/0022343310378914]. ACLED defines these events as incidents involving armed groups with political objectives as well as demonstrations, riots, and protests. With these totals in-hand, we needed to create a target to label our training data. Prior work has treated this as a binary classification task, asking whether or not conflict escalated in a given month, and we adopt this approach as well. [@10.1093/jeea/jvac025]
